**The statistical analysis of non-linear longitudinal data in biomedical research**

### 2021-01-05 Manuscripts/Manuscript_AM_v1.Rmd
Minor update, knit does not work on laptop

---
title: '**The statistical analysis of non-linear longitudinal data in biomedical research**'
author:
- Ariel Mundo^[Department of Biomedical Engineering, University of Arkansas, Fayetteville]
- Timothy J. Muldoon^[Department of Biomedical Engineering, University of Arkansas,
  Fayetteville]
- John R. Tipton^[Department of Mathematical Sciences, University of Arkansas, Fayetteville]
subtitle: _Beyond repeated measures ANOVA and Linear Mixed Models_
bibliography: refs.bib
csl: ieee.csl
css: style.css
output:

@@

A longitudinal study is defined as that which is designed to repeatedly measure a variable of interest in a group (or groups) of subjects. In biomedical research, this type of study is employed when the intention is to observe the evolution of the effect of a certain treatment across time, rather than analyzing it at a single time point (a cross-sectional study). Clinical studies of breast and neck cancer have used a longitudinal approach [@sio2016;@kamstra2015]. In the first case, weekly measurements of skin toxicities in  patients with radiation-induced dermatitis were taken for up to 8 weeks; whereas in the latter mouth opening was assessed at 6,12, 18, 24 and 36 months after radiotherapy (RT). Longitudinal studies have been also used to analyze tumor response [@roblyer2011;@tank2020;@pavlov2018;@demidov2018], antibody expression [@ritter2001;@roth2017], and cell metabolism [@jones2018;@skala2010].

>>>>

---
title: '**The statistical analysis of non-linear longitudinal data in biomedical research**'
author:
- Ariel Mundo^[Department of Biomedical Engineering, University of Arkansas, Fayetteville]
- Timothy J. Muldoon^[Department of Biomedical Engineering, University of Arkansas,
  Fayetteville]
- John R. Tipton^[Department of Mathematical Sciences, University of Arkansas, Fayetteville]
subtitle: _Beyond repeated measures ANOVA and Linear Mixed Models_
bibliography: refs.bib
csl: ieee.csl
css: style.css
fontfamily: Arial
  output:

@@

A longitudinal study is defined as that which is designed to repeatedly measure a variable of interest in a group (or groups) of subjects. In biomedical research, this type of study is employed when the intention is to observe the evolution of the effect of a certain treatment across time, rather than analyzing it at a single time point (a cross-sectional study). Clinical studies of breast and neck cancer have used a longitudinal approach [@sio2016;@kamstra2015]. In the first case, weekly measurements of skin toxicities in  patients with radiation-induced dermatitis were taken for up to 8 weeks; whereas in the latter mouth opening was assessed at 6,12, 18, 24 and 36 months after radiotherapy (RT). Longitudinal studies have beenn also used to analyze tumor response [@roblyer2011;@tank2020;@pavlov2018;@demidov2018], antibody expression [@ritter2001;@roth2017], and cell metabolism [@jones2018;@skala2010].
